---
title: "00_read_data"
author: "Ruth Ma"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output: html_document
---

```{r, install-packages}
# Install packages for gene analysis
if (!requireNamespace("Seurat", quietly = TRUE)) {
  install.packages("Seurat")
}
if (!requireNamespace("SeuratData", quietly = TRUE)) {
  install.packages("remotes")
  remotes::install_github('satijalab/seurat-data')
}
if (!requireNamespace("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2")
}
if (!requireNamespace("patchwork", quietly = TRUE)) {
  install.packages("patchwork")
}
if (!requireNamespace("dplyr", quietly = TRUE)) {
  install.packages("dplyr")
}
# Install 'here' package
if (!requireNamespace("here", quietly = TRUE)) {
  install.packages("here")
}
if (!requireNamespace("gtsummary", quietly = TRUE)) {
  install.packages("gtsummary")
}
```

```{r, import-package}
# Import packages for gene analysis
library(Seurat)
library(SeuratData)
library(ggplot2)
library(patchwork)
library(dplyr)
# Import 'here' package
library(here)
```

Get the current working directory
```{r}
here::i_am("code/00_read_data.Rmd")
```

# Read in the dataset
## Method 1: Read in the dataset online
```{r, install-data}
options(timeout = 600)
if (!(AvailableData()['stxBrain','Installed']==TRUE)) {
  InstallData("stxBrain")
}
```

```{r, read-data-from-web}
rna_data <- LoadData("stxBrain", type = "posterior2")
```

## Method 2: Read in local dataset
```{r, read-data-from-local, eval=FALSE}
data_dir <- here::here('data')
file_name = 'V1_Mouse_Brain_Sagittal_Posterior_Section_2_filtered_feature_bc_matrix.h5'
rna_data <- Load10X_Spatial(data.dir = data_dir, filename = file_name)
```

We will calculate mitochondrial quality control metrics.  If cells exhibit mitochondrial counts exceeding either 5% or 10%, it could indicate potential cell damage, and we will filter these cells out.
```{r}
rna_data[["percent.mt"]] <- PercentageFeatureSet(rna_data, pattern = "^mt-")
```

We convert the h5 data type to rds format to make it more accessible for those who may not be familiar with h5.
```{r}
saveRDS(
  rna_data, 
  file = here::here("derived_data/rna_data.rds")
)
```